---
title: "Scripts to generate Figure 2"
output:
  html_document:
    df_print: paged
---

### MDS plot for illumina metagenomics data

Load generic libraries
```{r message=FALSE, warning=FALSE}
source('configuration.r')
```

Load specific libraries
```{r message=FALSE, warning=FALSE}
library(vegan)
library(ComplexHeatmap)
library(RColorBrewer)
library(foreach)
library(forcats)
library(magrittr)
library(ggpubr)
```

Preprocess data
```{r}
meta <- read.table('../metadata/illumina_metadata.txt', head=TRUE, row.names=2)
##meta <- meta[which(meta$timept !=3), ]

dat <- read.table('../output_tables/metagenomics.metaphlan2.table.filtered.g', head=TRUE, row.names=1)
dat[dat < 0.1] <- 0
dat <- dat[rowSums(dat)>0, ]

meta.fil <- meta %>% select(Library,Room_type, Sample_type, Cubicle_room, timept)
fil <- filter(meta.fil, Room_type != 'GIS' & timept == 1)  ## remove samples from GIS, take only the first time point
dat.fil <- dat[, as.character(fil$Library)]

## do MDS
dist.mat <- vegdist(t(dat.fil))
cmds <- cmdscale(dist.mat, k=3, eig=TRUE)
eigen <- cmds$eig / sum(cmds$eig) * 100

dat.merged <- (merge(cmds$points, meta.fil, by.x=0, by.y=0, all.x=TRUE)) %>% 
  mutate(Sample_type=str_replace(Sample_type, '_', ' ')) %>% 
  mutate(Sample_type=str_replace(Sample_type, 'Door handle-interior', 'Door Handle'))
dat.merged$Sample_type <- fct_relevel(factor(dat.merged$Sample_type), 'Sink Trap')
```

PCoA1 and PCoA2
```{r fig.height=5, fig.width=9}
## PCoA
ggplot(dat.merged, aes(x=V1, y=V2, col=Sample_type, fill=Sample_type, shape=Sample_type), lwd=2) +
  geom_density_2d(aes(x=V1, y=V2),  inherit.aes = FALSE, col='grey', lwd=1) + 
	geom_point(size=2.5) +
	labs(x=paste0('MDS1 (',round(eigen[1], 1),'%)'),
       y=paste0('MDS2 (',round(eigen[2], 1),'%)')) +
	scale_color_manual(values=c(pal_npg(c("nrc"))(10)[c(1,5,7,2,3,10,4)], 'grey')) +  
  scale_fill_manual(values=c(pal_npg(c("nrc"))(10)[c(1,5,7,2,3,10,4)], 'grey')) +
  scale_shape_manual(values=c(8,9,24,25, 21,22,23)) + 
  theme(legend.title = element_blank()) 
ggsave('../plots/f2a_mds_genus.svg', width =9, height = 5)
```

### Analysis with the handling controls

```{r fig.height=6, fig.width=8}
dat.ori <- read.table('../tables/metagenomics.metaphlan2.table.g', head=TRUE, row.names=1)
dat.ori[dat.ori < 0.1] <- 0
dat.ori <- dat.ori[rowSums(dat.ori)>0, ]
handling.controls <- filter(meta.fil, Sample_type=='Handling_Control') %>% pull(Library) %>% as.character()
fil.tp1_tp2 <- filter(meta.fil, Room_type != 'GIS' & timept %in% c(1,2))
dat.tp1_tp2 <- dat.ori[,  fil.tp1_tp2$Library]

qubit <- read.table('../metadata/metagenomics_qubit.txt',head=T,check.names = F, sep='\t', stringsAsFactors = F) %>% 
  mutate(cat=cut(`DNA Amt`, c(0.5, 1,3,Inf))) %>% 
  filter(`Library ID` %in% fil.tp1_tp2$Library)

dat.fil.control <- cbind(dat.ori[, handling.controls], dat.tp1_tp2)
dist.mat <- as.matrix(vegdist(t(dat.fil.control)))

g <- c(rep('Handling control', length(handling.controls)), ifelse(with(fil.tp1_tp2, Sample_type == "Aerator" | Sample_type == "Sink_Trap"), 'CTB', 'CTA'))

## HC vs HC
tmp <- dist.mat[g == 'Handling control', g == 'Handling control']
d1 <- data.frame(bray=tmp[lower.tri(tmp)], cat='Control')

d2 <- foreach(c=levels(qubit$cat), .combine = rbind) %do% {
  tmp <- filter(qubit, cat==c) %>% pull(`Library ID`)
  d <- data.frame(bray=c(dist.mat[tmp, g=='Handling control']), cat=c, stringsAsFactors = F)
  d
}


# ## HC vs CTA
# tmp <- dist.mat[g == 'Handling control', g == 'CTA']
# d2 <- data.frame(bray=c(tmp), cat='control vs. CTA')
# ## HC vs CTB
# tmp <- dist.mat[g == 'Handling control', g == 'CTB']
# d3 <- data.frame(bray=c(tmp), cat='control vs. CTB')

plot.dat <- rbind(d1, d2)
  
ggplot(plot.dat, aes(x=cat, y=bray, fill=cat)) + 
  #geom_bar(data=rbind( data.frame(cat='Control', n=4), count(qubit, cat)), aes(x=cat, y=n/200), alpha=0.5, stat='identity') + 
  geom_boxplot(lwd=1) + 
  theme(axis.text.x = element_text(angle=40, hjust=1), axis.title.y.right = element_text(margin=margin(0,0,0,15))) + 
  scale_y_continuous(breaks=c(0, 0.25, 0.5, 0.75, 1), limits=c(0,1.5)) +#, sec.axis = sec_axis(~.*200, name = "Number of samples")) + 
  scale_fill_aaas(guide=F) + 
  labs(x='DNA amount (ng)', y='Bray-Curtis distance') + 
  stat_compare_means(comparisons = list(c("Control", "(0.5,1]"), 
                                        c("Control", "(1,3]"),
                                        c("Control", "(3,Inf]")
                                        ),
                     label = "p.signif")

ggsave('../plots/supnote1a_handling_control_vs_samples.png', width =5, height = 5)
```



### Differential abundance taxa between wet and dry

Genus level

```{r }
rowMedians <- function(x) apply(x, 1, median) ## aux function
wet.sel <- with(fil, Sample_type == "Aerator" | Sample_type == "Sink_Trap")

g <- factor(wet.sel, labels = c("D","W"))

## Select common and abundant species in Wet/Dry 
medianDry <- rowMedians(dat.fil[, !wet.sel])
medianWet <- rowMedians(dat.fil[, wet.sel])
data1 <- dat.fil[which(medianDry > 0.5 | medianWet > 0.5),]
data1 <- data1[-grep("unclassified",rownames(data1)),]

rownames(data1) %<>% str_replace_all(c("g__"="", "_noname"="","Propionibacterium"="Cutibacterium" ))

## wilcoxon test
p.values <- sapply(seq(dim(data1)[1]), function(x) wilcox.test(as.numeric(data1[x,(g=="D")]),as.numeric(data1[x,g=="W"]))$p.value)
q.values <- p.adjust(p.values, method="fdr")

#filter data to plot heatmap
data2 <- data1[which(q.values<0.01),]
data2dry <- data2[rowMedians(data2[,!wet.sel])>rowMedians(data2[,wet.sel]),]
data2wet <- data2[rowMedians(data2[,!wet.sel])<rowMedians(data2[,wet.sel]),]

data2Vdry <- cbind(rowMedians(data2dry[,!wet.sel]),rowMedians(data2dry[,wet.sel]))
data2Vwet <- cbind(rowMedians(data2wet[,!wet.sel]),rowMedians(data2wet[,wet.sel]))
data2all <- rbind(data2Vdry,data2Vwet)
rownames(data2all) <- rownames(rbind(data2dry,data2wet))
colnames(data2all) <- c('CTA', 'CTB')
```

Validation and filtering with time point 3

```{r message=FALSE, warning=FALSE, fig.width=2.5, fig.height=4}
genus_comp <- rownames(data2all)
fil.tp3 <- filter(meta.fil, Room_type != 'GIS' & timept == 3)  ## remove samples from GIS, take only the third time point
wet.sel.tp3 <- with(fil.tp3, Sample_type == "Aerator" | Sample_type == "Sink_Trap")
g.tp3 <- factor(wet.sel.tp3, labels = c("D","W"))
dat.fil.tp3 <- dat[, as.character(fil.tp3$Library)]
rownames(dat.fil.tp3) %<>% str_replace_all(c("g__"="", "_noname"="","Propionibacterium"="Cutibacterium" ))
data1 <- dat.fil.tp3[genus_comp, ]
## filter1: p-value < 0.01
p.values <- sapply(seq(dim(data1)[1]), function(x) wilcox.test(as.numeric(data1[x,(g.tp3=="D")]),as.numeric(data1[x,g.tp3=="W"]))$p.value) < 0.01
## filter2: trend consistent
trend <- c(rowMedians(data1[rownames(data2dry),g.tp3=='D']) > rowMedians(data1[rownames(data2dry), g.tp3=='W']), 
  rowMedians(data1[rownames(data2wet),g.tp3=='D']) < rowMedians(data1[rownames(data2wet), g.tp3=='W']))

#plot heatmap
p <- Heatmap(log(data2all[p.values & trend, ]+1,2), col=(colorRampPalette(brewer.pal(9,'Blues'))(60)),
        rect_gp = gpar(col = "black", lty = 1, lwd = 1),
        heatmap_legend_param = list(title='', legend_height=unit(7, "cm"), labels_gp=gpar(fontsize=15), at=0:5),
        row_names_gp=gpar(fontface = "bold.italic"), column_names_gp = gpar(fontface = "bold"),
 border="black",cluster_rows = F, cluster_columns = F)
p
pdf("../plots/fig2bi_dry_wet_genus_comparison.pdf", height = 4, width = 2.5)
p
dev.off()
```

Species level

```{r fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
dat.species <- read.table('../output_tables/metagenomics.metaphlan2.table.filtered.s', head=TRUE, row.names=1)
dat.species[dat.species < 0.1] <- 0
dat.species <- dat.species[rowSums(dat.species)>0, ] 
rownames(dat.species) %<>% str_replace_all(c("s__"="", "_noname"="", "_"=" "))

dat.species.tp1 <- dat.species[, as.character(fil$Library)]
dat.species.tp3 <- dat.species[, as.character(fil.tp3$Library)]

medianDry <- rowMedians(dat.species.tp1[!wet.sel])
medianWet <- rowMedians(dat.species.tp1[wet.sel])
data1 <- dat.species.tp1[which(medianDry > 0.5 | medianWet > 0.5),]
data1 <- data1[-grep("unclassified",rownames(data1)),]

## wilcoxon test for the 1st time point
p.values <-  apply(data1,1, function(x) wilcox.test(x[g=="D"], x[g=="W"])$p.value)
q.values <- p.adjust(p.values, method="fdr")
data2 <- data1 %>% rownames_to_column("Species") %>% 
  filter(q.values<0.01)

## validation using the 3rd time point
data1.tp3 <- dat.species.tp3[data2$Species,]
p.values.tp3 <- apply(data1.tp3,1, function(x) wilcox.test(x[g.tp3=="D"], x[g.tp3=="W"])$p.value)

data2 %<>% filter(p.values.tp3<0.01)

## trend
data2dry <- data2[rowMedians(data2[,-1][,!wet.sel])>rowMedians(data2[,-1][,wet.sel]),]
data2wet <- data2[rowMedians(data2[,-1][,!wet.sel])<rowMedians(data2[,-1][,wet.sel]),]
## trend validation using the 3rd time point -- this doesn't remove any species
data2dry <- data2dry[rowMedians(data1.tp3[data2dry$Species,!wet.sel.tp3])>rowMedians(data1.tp3[data2dry$Species,wet.sel.tp3]), ]
data2wet <- data2wet[rowMedians(data1.tp3[data2wet$Species,!wet.sel.tp3])<rowMedians(data1.tp3[data2wet$Species,wet.sel.tp3]), ]

data2dryP <- filter(data2dry,Species=="Propionibacterium acnes")
data2dry <- filter(data2dry,Species!="Propionibacterium acnes")

anno <- data.frame(variable=colnames(dat.species.tp1), group=g)
data2dryP <- melt(data2dryP,id=c("Species")) %>% merge(anno)
data2dry <- melt(data2dry,id=c("Species")) %>% merge(anno)
data2wet <- melt(data2wet,id=c("Species")) %>% merge(anno)

### Enrich in dry
p1 <- ggplot(data = data2dry, aes(x=Species,y=value,fill=group, grp=group)) +
	geom_boxplot(width=1,outlier.shape=NA, lwd=1) +
  coord_cartesian(ylim=c(0,8)) +
  scale_fill_manual(values=pal_npg(c("nrc"))(10)[c(7,1)]) + 
  labs(y=NULL, x=NULL) +
	theme(axis.text.x = element_text(angle=40, hjust=1, size=10), legend.position = 'left', legend.title = element_blank()) 
### Enrich in dry (C. acnes)
p2 <- ggplot(data = data2dryP, aes(x=Species,y=value,fill=group, grp=group)) +
	geom_boxplot(width=1,outlier.shape=NA, lwd=1) +
  labs(y=NULL, x=NULL) +
  scale_fill_manual(values=pal_npg(c("nrc"))(10)[c(7,1)]) + 
	theme(axis.text.x = element_text(angle=40, hjust=1, size=10),legend.position = 'none') 
### Enrich in wet
p3 <- ggplot(data = data2wet, aes(x=Species,y=value,fill=group, grp=group)) +
	geom_boxplot(width=1,outlier.shape=NA, lwd=1) +
  scale_fill_manual(values=pal_npg(c("nrc"))(10)[c(7,1)]) + 
  coord_cartesian(ylim=c(0,50)) +
  labs(y=NULL, x=NULL) +
	theme(axis.text.x = element_text(angle=40, hjust=1, size=10), legend.position = 'none')

cowplot::plot_grid(p1, p2, p3, nrow=1, rel_widths=c(12,5, 8),align ='h')
ggsave('../plots/fig2bii_dry_wet_species_comparison.svg', height = 7, width = 10)
```

### Spatial-temporal analysis

```{r fig.height=8, fig.width=16}
meta.fil <- meta %>% select(Sample_ID, Room_type, Sample_type, timept, bed_number, Cubicle_room, Library)
##group
meta.fil[which(is.na(meta.fil$bed_number)),]$bed_number <- 1
dat_t <- data.frame((t(dat)))%>% rownames_to_column('Library')
meta.select <- meta.fil %>%
  filter(timept %in% c(1,2) ) %>%  ## remove 3rd timepoint and controls
  group_by(Room_type,Cubicle_room,Sample_ID,Sample_type,timept,Library)%>% 
  summarise()%>% 
  merge(dat_t,by="Library") #,all.y=T)

dat_select=data.frame(select(meta.select,-Sample_type, -timept, -Sample_ID, -Room_type, -Cubicle_room),row.names = "Library")
metadat_select=data.frame(select(meta.select, Sample_type, timept, Library, Sample_ID, Room_type, Cubicle_room),row.names = "Library")

## PCoA
dist.mat <- vegdist(dat_select)
cmds <- cmdscale(dist.mat, eig=TRUE)
eigen <- cmds$eig / sum(cmds$eig) * 100
dat.merged <- (merge(cmds$points, metadat_select, by.x=0, by.y=0, all.x=TRUE))

dat.merged <- filter(dat.merged, !is.na(Room_type), Room_type!="GIS") 
levels(dat.merged$Sample_type) <- stringr::str_replace_all(levels(dat.merged$Sample_type), c('_'=' ','-'=' ','Door handle interior'='Door Handle'))
levels(dat.merged$Room_type) <- stringr::str_replace_all(levels(dat.merged$Room_type), c('_'=' ', 'Non-cohort'='Non-controlled'))
dat.merged$Sample_type <- relevel(factor(dat.merged$Sample_type), 'Sink Trap')

## links:
from <- dat.merged %>% filter(timept==1) %>% arrange(Sample_ID) %>%
  select(V1,V2,Sample_ID) 
to <- dat.merged %>% filter(timept==2) %>% arrange(Sample_ID)%>%
  select(V1,V2,Sample_ID)
arrows <- (merge(from, to, by='Sample_ID'))

plot.dat <- merge(dat.merged, arrows, by='Sample_ID')
plot.dat$Time_point <- as.factor(plot.dat$timept)

ggplot(plot.dat, aes(x=V1, y=V2, col=Sample_type, shape=Time_point), lwd=2) +
  geom_curve(data=plot.dat, aes(x=V1.x,y=V2.x,xend=V1.y, yend=V2.y),
             arrow = arrow(length = unit(0.02, "npc")), lwd=1, alpha=0.5,
             inherit.aes = FALSE) +
  geom_point(size=3, alpha=0.9) + 
  coord_cartesian(ylim = c(-0.5, 0.5), xlim = c(-0.7,0.7)) +
  labs(x=paste0('MDS1 (',round(eigen[1], 1),'%)'),
       y=paste0('MDS2 (',round(eigen[2], 1),'%)')) +
  scale_shape_manual(values=c(17, 19)) + 
  scale_color_manual(values=pal_npg(c("nrc"))(10)[c(1,5,7,2,3,10,4)]) +
  facet_wrap(~Sample_type,dir="v") + 
  guides(color=guide_legend(title='Sample type'), shape=guide_legend('Time point'))

ggsave('../plots/fig2c_pcoa_spatial_temporal.svg', height = 8, width = 17)
```

Test if the two CTs have different stability
```{r}
idx <- arrange(plot.dat, Sample_ID, timept) %>% 
  select(Row.names, Sample_type, timept)

mat <- as.matrix(dist.mat)
d.pair <- foreach(r=seq(1,nrow(idx),2), .combine=rbind) %do%{
  data.frame(d=mat[idx[r, 1], idx[r+1, 1]], type=idx[r, 2])
}

wilcox.test(d.pair$d[d.pair$type %in% c("Aerator", "Sink Trap")], d.pair$d[!d.pair$type %in% c("Aerator", "Sink Trap")])
```

### Distance between samples

```{r fig.height=8, fig.width=4}
meta.fil <- meta %>% select(Library, Room_type, Sample_type, timept, Cubicle_room, bed_number)
fil <- filter(meta.fil, Room_type != 'GIS' & Room_type != 'Handling controls' & timept == 1) %>% 
  filter((Sample_type == "Bed_Rail" | Sample_type == "Bedside_Locker" | Sample_type == "Cardiac_Table")) %>% 
  mutate(MDRO=ifelse(Room_type == "MDRO_cubicles", "MDRO_","")) %>% unite(MDRO,Cubicle_room,col="Cubicle_room",sep="")
fil[which(is.na(fil$bed_number)),]$bed_number <- 1

dat.fil <- dat[, as.character(fil$Library)]

braymatrix <- as.matrix(vegdist(t(dat.fil),method="bray"))

grouped_ids <- sapply((group_by(fil,Cubicle_room,bed_number) %>%
                         summarise(id=str_c(Library, collapse=',')))$id, str_split, pattern=',')
grouped_ids_sameroom <- sapply((group_by(fil,Cubicle_room) %>%
                                  summarise(id=str_c(Library, collapse=',')))$id, str_split, pattern=',')
grouped_ids_samemat <- sapply((group_by(fil,Sample_type) %>%
                                 summarise(id=str_c(Library, collapse=',')))$id, str_split, pattern=',')

aux <- function(x){
  tmp = braymatrix[x, x]
	tmp[upper.tri(tmp)]
}

same_bed=c(sapply(grouped_ids, aux))
braymatrix.copy <- braymatrix

for(x in grouped_ids_sameroom){
  braymatrix.copy[x, x] <- NA
}
for(x in grouped_ids_samemat){
  braymatrix.copy[x, x] <- NA
}
diff_rooms <- braymatrix.copy[upper.tri(braymatrix.copy)]
diff_rooms <- diff_rooms[!is.na(diff_rooms)]

braymatrix.copy <- braymatrix
for(x in grouped_ids){
  braymatrix.copy[x, x] <- NA
}
for(x in grouped_ids_samemat){
  braymatrix.copy[x, x] <- NA
}

aux_room <- function(x){
  tmp = braymatrix.copy[x, x]
	tmp[upper.tri(tmp)]
}

same_room <- unlist(sapply(grouped_ids_sameroom, aux_room),use.names=F)
same_room <- same_room[!is.na(same_room)]

fil=filter(meta.fil,Room_type != 'GIS' & Room_type != 'Negative controls' & timept < 3) %>% 
  filter((Sample_type == "Bed_Rail" | Sample_type == "Bedside_Locker" | Sample_type == "Cardiac_Table")) %>% 
  mutate(MDRO=ifelse(Room_type == "MDRO_cubicles", "MDRO_","")) %>% unite(MDRO,Cubicle_room,col="Cubicle_room",sep="")
fil[which(is.na(fil$bed_number)),]$bed_number=1
timedat.fil <- dat[, as.character(fil$Library)]

braymatrix=as.matrix(vegdist(t(timedat.fil),method="bray"))
grouped_ids_timepoints <- sapply((group_by(fil,Cubicle_room,bed_number,timept) %>%
                                    summarise(id=str_c(Library, collapse=',')))$id, str_split, pattern=',')

timepoint_tmp=list()
for (i in (seq(from = 1, to = length(grouped_ids_timepoints), by =2))) {
	timepoint_tmp[[(i+1)/2]]=braymatrix[grouped_ids_timepoints[[i]], grouped_ids_timepoints[[i+1]]]
}

for (i in (1:length(timepoint_tmp))){
  timepoint_tmp[[i]][1,1]<-NA  
	timepoint_tmp[[i]][2,2]<-NA  
	timepoint_tmp[[i]][3,3]<-NA  
}

timepoint_mat=unlist(timepoint_tmp)
timepoint_mat=timepoint_mat[!is.na(timepoint_mat)]

rbind(
  data.frame(dist=same_bed, lab='Within bed (T1)'),
  data.frame(dist=timepoint_mat, lab='Across time'),
  data.frame(dist=same_room, lab='Within ward (T1)'),
  data.frame(dist=diff_rooms, lab='Across wards (T1)')
) %>% 
  ggplot(aes(x=lab, y=dist, fill=lab)) + 
  geom_boxplot(lwd=1) + 
  scale_fill_npg(guide=F) + 
  labs(x=NULL, y=NULL) +
  scale_y_continuous(limits=c(0,1)) +
  theme(axis.text.x = element_text(angle=45, hjust=1))
ggsave('../plots/sup3a_distance_spatial_temporal.svg', height = 8, width = 4)
## ad hoc analysis
#median(same_bed) 
#median(timepoint_mat)
#median(same_room)
#median(diff_rooms)

wilcox.test(same_bed,timepoint_mat)
wilcox.test(same_bed,same_room)
wilcox.test(same_bed,diff_rooms)
wilcox.test(same_room, timepoint_mat)
```

### Species differentially abundant in dry environments

```{r fig.height=6, fig.width=8}
mat.sel <- rownames_to_column(dat.species, "species") %>% 
  filter(grepl("Burkholderia", species) | 
               species %in% c("Acinetobacter baumannii","Candida albicans",
                         "Clostridium difficile", "Clostridium sordellii", 
                         "Klebsiella pneumoniae", "Klebsiella oxytoca",
                         "Escherichia coli", "Staphylococcus aureus",
                         "Pseudomonas aeruginosa","Mycobacterium abscessus",
                         "Mycobacterium tuberculosis","Enterococcus faecalis",
                         "Enterococcus faecium", "Staphylococcus epidermidis")) %>% 
  column_to_rownames("species")

df <- merge(meta, data.frame(t(mat.sel), Site=colnames(mat.sel)), by=0,all.y=T) %>% 
  filter(!Room_type %in% c('GIS','Handling_Control', 'Mock', 'Negative_Control') )%>% 
  filter(timept %in% c(1,2)) %>% 
  select(-Site,-Sample_ID,-Cubicle_room,-Material,-Human_Traffic,-timept,-bed_number) %>% 
  column_to_rownames("Row.names")
colnames(df)=gsub("g__","",colnames(df))
colnames(df)=gsub("s__","",colnames(df))
df[,"Sample_type"]=gsub("_"," ",df[,"Sample_type"])
df[(which(df[,"Sample_type"]=="Door handle-interior")),"Sample_type"] = "Door Handle"

mutate(df, Env=ifelse((Sample_type=="Aerator" | Sample_type=="Sink Trap"),"Wet","Dry")) %>%
  select(-Sample_type, -Library) %>% 
  filter(Env=="Dry") %>% 
  select(-Env) -> all.dat.dry.fil

### Test for significance
dry.species <- group_by(all.dat.dry.fil,Room_type) %>%
  summarise_all(median) %>% 
  select_if(~is.numeric(.)) %>% 
  select_if(~max(.)>=0.5) %>% 
  colnames()
pvalueD=p.adjust(sapply(dry.species, function(x) kruskal.test(as.formula(paste0(x, "~Room_type")),data=all.dat.dry.fil)$p.value),method="fdr")

all.dat.dry.fil <- melt(all.dat.dry.fil,id="Room_type") %>% 
  filter(variable%in%names(pvalueD[which(pvalueD<0.01)])) %>% 
  mutate(variable=str_replace_all(variable,"s__","")) %>% 
  mutate(variable=str_replace_all(variable,"_"," ")) %>%
  mutate(Room_type=str_replace_all(Room_type,c('cubicles'='wards', 'Non-cohort'='Standard',"_"=" "))) %>% 
  mutate(variable=str_replace(variable, '[a-z]+ ', '. '))


ggplot(data = all.dat.dry.fil, aes(x=variable,y=value,fill=Room_type)) +
  geom_boxplot(outlier.shape = NA, lwd=1) +
  theme(axis.text.x = element_text(angle=40, hjust=1), plot.margin = unit(c(0, 0, 0, 0.5), "cm"),
        legend.position = 'top', legend.title = element_blank(), legend.key.size =unit(2,'line'))+
  guides(fill = guide_legend(ncol = 1)) + 
  scale_fill_npg() + 
  labs(y="Relative Abundance", x=NULL) + 
  coord_cartesian(ylim =c(0,5))

ggsave('../plots/fig2e_species_in_dry_environments.svg', height = 7, width = 4)
```



```{r echo=FALSE, eval=FALSE}
### PERMANOVA test
df.fil <- df[rowSums(df[,-c(1:3)]) > 0, -c(1,3)]
df.fil %>% filter(!Sample_type %in% c('Aerator','Sink Trap')) -> dry.dat

adonis2(df.fil[,-1] ~ site, data = data.frame(site=df.fil$Sample_type))
adonis2(dry.dat[,-1] ~ site, data = data.frame(site=dry.dat$Sample_type))
```

### Session informaton
```{r}
sessionInfo()
```